Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-10 doc/mc-paper.tex

\authorrunning{R. E. Bryant, et al} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

@@

\begin{abstract}
  Knowledge compilers convert Boolean formulas into a form 
  that enables efficient evaluation of properties of the formula,
  including
  weighted and unweighted model counting.
 Existing knowledge compilers, however,
provide no guarantee that the generated representation accurately
describes the original formula.

We present the Partitioned Operation Graph (POG) representation that
generalizes existing knowledge compiler representations, and a proof
framework that can prove the equivalence of a POG representation with
a conjunctive normal form (CNF) Boolean formula.

We have developed a program that can generate POG representations and
equivalence proofs from the d-DNNF graphs generated by a leading
knowledge compiler.  We also present two proof
checkers---a fast checker that can handle the large proofs generated,
and a formally verified checker written in the \lean{} proof system.
Our generation and checking tool chain scales to all but the largest formulas from a
recent model counting competition.  
\end{abstract}

>>>>

\authorrunning{R. E. Bryant et al.} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

@@

\begin{abstract}
\emph{Knowledge compilation} is the process of converting a Boolean
formula into a form that makes it possible to calculate associated
values efficiently, such as the weighted or unweighted model
count. Existing knowledge compilers, however,
provide no guarantee that the representation they generate
is logically equivalent to the original formula, and therefore they cannot validate the generated counts.

We present \emph{Partitioned Operation Graphs} (POGs), a generalization
of representations commonly used in knowledge compilation.
We have designed a proof framework that can prove the equivalence of a
POG representation with a Boolean formula in conjunctive normal form (CNF),
the standard input format for SAT solvers and model counting tools.

We have developed a program that can generate POG representations
from the d-DNNF graphs generated by the state-of-the-art knowledge
compiler D4, as well as independently checkable proofs that the
output POGs are equivalent to the input CNF formulas.
We have implemented an efficient checker
that can handle very large proofs. Our tool
chain for generating and verifying POGs scales to all but the largest d-DNNF graphs
generated by D4 for formulas
from a recent model counting competition. Moreover, we
present a checker that has been formally verified
in the Lean~4 proof assistant, thereby creating the
first formally verified toolchain for weighted and unweighted model
counting.

%  Knowledge compilers convert Boolean formulas into a form 
%  that enables efficient evaluation of properties of the formula,
%  including
%  weighted and unweighted model counting.
% Existing knowledge compilers, however,
%provide no guarantee that the generated representation accurately
%describes the original formula.
%
%We present the Partitioned Operation Graph (POG) representation that
%generalizes existing knowledge compiler representations, and a proof
%framework that can prove the equivalence of a POG representation with
%a conjunctive normal form (CNF) Boolean formula.
%
%We have developed a program that can generate POG representations and
%equivalence proofs from the d-DNNF graphs generated by a leading
%knowledge compiler.  We also present two proof
%checkers---a fast checker that can handle the large proofs generated,
%and a formally verified checker written in the \lean{} proof system.
%Our generation and checking tool chain scales to all but the largest formulas from a
%recent model counting competition.  
\end{abstract}
